---
title: "DataPreparation"
output: html_document
date: "2024-09-20"
---

## Introduction
After the data exploration, the data will be prepared for the analysis. First, all the NAs are removed from the datasets, after which the three datasets will be merged together. At last, a new variable is created that show the length of the series, which will be the independent variable in the analysis. 


## Preparation
### Load packages
```{r, echo=FALSE}
library(tidyr)
```

### Remove NAs from title_basics dataset and show the new summary statistics
```{r, echo=FALSE}
title_basics_no_NAs <- drop_na(title_basics)
View(title_basics_no_NAs)
summary(title_basics_no_NAs)
```

### Remove NAs from title_episode dataset and show the new summary statistics
```{r, echo=FALSE}
title_episode_no_NAs <- drop_na(title_episode)
View(title_episode_no_NAs)
summary(title_episode_no_NAs) 
```

### Remove NAs from title_ratings dataset and show the new summary statistics
```{r, echo=FALSE}
title_ratings_no_NAs <- drop_na(title_ratings)
View(title_ratings_no_NAs)
summary(title_ratings_no_NAs) 
```

### Merge the three datsets together
```{r, echo=FALSE}
merged_data <- title_basics_no_NAs %>%
  inner_join(title_ratings_no_NAs, by = "tconst") %>%
  inner_join(title_episode_no_NAs, by = c("tconst" = "parentTconst"))
```

### Create new variabel: series_length
A new variable is created that shows how many years a series has run, called series_length. This variable will be the independent variable in our analysis, and is created by deducting the first year that the series ran from the last year. 
```{r, echo=FALSE}
merged_data <- merged_data %>%
  mutate(series_lengths = endYear - startYear)
```