---
title: "Plots for regression analysis"

author: "team 03"
date: "2024-09-27"
output: html_document
---

## Assumptions of Linear Regression

The assumptions of the linear regression models by generating diagnostic plots.

```{r setup, include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
library(ggplot2)
library(car)
```

### Model 1: total_years

```{r model1, include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
model1 <- lm(averageRating ~ total_years + log_numVotes, data = cleaned_data)
summary(model1)
```


#### Residuals vs. Fitted Plot

```{r residuals plot,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}

plot(model1, which = 1)

```


#### Q-Q Plot for normality check

```{r qqplot,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
plot(model1, which = 2)

```


#### Scale-Location Plot for homoscedasticity check

```{r scale-loc plot,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
plot(model1, which = 3)
```


#### Residuals vs. Leverage Plot to detect influential points

```{r residual leverage,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
plot(model1, which = 5)
```




### Model 2: log_episode_count

```{r model2, include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
model2 <- lm(averageRating ~ log_episode_count + log_numVotes, data = cleaned_data)
summary(model2)
```


#### Residuals vs. Fitted Plot

```{r residuals plot,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}

plot(model2, which = 1)

```


#### Q-Q Plot for normality check

```{r qqplot,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
plot(model2, which = 2)

```


#### Scale-Location Plot for homoscedasticity check

```{r scale-loc plot,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
plot(model2, which = 3)
```


#### Residuals vs. Leverage Plot to detect influential points

```{r leverage,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
plot(model2, which = 5)
```




