---
title: "Data Exploration and Preparation"
author: "team 03"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message= FALSE)
```


### Install necessary pacakges for data exploration

The following libraries are required for this analysis: 

-**readr**: For reading data files.
-**tidyverse**: For data manipulation and analysis.
-**dplyr**: For data wrangling and manipulation.
-**ggplot2**: For data visualization.

```{r install packages,include=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Introduction

This report provides the initial data exploration of the data sets. Our goal is to provide the structure of our data set,summarize key variables,and provide some basic exploratory visualizations. The data set contains information about TV series, including customer ratings, number of episodes and series duration.
 
## Overview of the Dataset and Variables

The dataset used in the project contains information about TV series and episodes from IMD Since we are interested in understanding the relationship between the length of the TV series and it's average customer ratings, we focus on the three datasets listed and described below.

-**title.episode.tsv.gz**: Contains information about individual episodes of TV series.
-**title.basics.tsv.gz**: Contains information about TV series titles,the start year,genres,etc.
-**title.ratings.tsv.gz**: Contains user ratings for titles.


## Description of Variables

#### title.episode.tsv.gz

```{r variables-episode,echo=FALSE}
variables_episode <- data.frame(
  Variables = c("`tconst`","`parentTconst`","`seasonNumber`", "`episodeNumber`"),
  Description = c(
    "An alphanumeric identifier unique to each episode.",
    "An alphanumeric identifier for the parent TV series of the episode.It links the episode to the overall series.",
    "The season number that the episode belongs to within the TV series.",
    "The specific episode number of the `tconst` in the TV series."
                  ),
  Class = c("String","String","Interger","Interger")
)

# Display the table using knitr::kable()
knitr::kable(variables_episode,col.names = c("Variables","Description","Class"))
```

#### title.basics.tsv.gz

```{r variables-basics,echo=FALSE}
# Create a data frame for title.basics.tsv.gz
variables_basics <- data.frame(
  Variables = c("`startYear`", "`endYear`"),
  Description = c(
    "The year the series began.",
    "TV Series end year."
                 ),
  Class = c("Integer", "Integer")
)

# Display the table using knitr::kable()
knitr::kable(variables_basics, col.names = c("Variables", "Description","Class"))

```

#### title.ratings.tsv.gz

```{r variables-ratings, echo=FALSE}
# Create a data frame for title.ratings.tsv.gz
variables_ratings <- data.frame(
  Variables = c("`tconst`", "`averageRating`"),
  Description = c(
    "Alphanumeric unique identifier of the title.",
    "Weighted average of all individual user ratings."
  ),
  Class = c("String", "Numeric")
)

# Display the table using knitr::kable()
knitr::kable(variables_ratings, col.names = c("Variables", "Description", "Class"))

```


### Loading the Datasets

We will programmatically load the following datasets from IMDb:

- **title.basics.tsv.gz**: Contains information about movie and TV series titles, their start year, genres, etc.
- **title.episode.tsv.gz**: Contains information about individual episodes of TV series.
- **title.ratings.tsv.gz**: Contains user ratings for titles.

```{r load-datasets, echo=FALSE, warning=FALSE, message=FALSE}

# URLs to the IMDb datasets
urls <- c(
  "https://datasets.imdbws.com/title.basics.tsv.gz",
  "https://datasets.imdbws.com/title.episode.tsv.gz",
  "https://datasets.imdbws.com/title.ratings.tsv.gz"
)

# Load the datasets
datasets <- lapply(urls, read_delim, delim = '\t', na = '\\N')

# Access individual datasets
title_basics <- datasets[[1]]
title_episode <- datasets[[2]]
title_ratings <- datasets[[3]]

```

### Initial Data Exploration

#### Summary Statistics 
##### Get a quick overview of the data by using the code summary()

Let's take a look at the basic summary statistics for each dataset.

```{r summary-stats, echo=FALSE, warning=FALSE, message=FALSE}
# Summary statistics for each dataset
summary(title_basics)
summary(title_episode)
summary(title_ratings)
```

#### Specific summary statistics
##### Summary statistics of title_episode

```{r specific-summary-stats, echo=FALSE, warning=FALSE, message=FALSE}

title_episode_summary <- title_episode %>%
  summarise(num_rows=n(),
            num_cols=ncol(.),
            missing_values_tconst=sum(is.na(tconst)),
            missing_values_parentTconst=sum(is.na(parentTconst)),
            missing_values_seasonNumber=sum(is.na(seasonNumber)),
            missing_values_episodeNumber=sum(is.na(episodeNumber)) 
  )%>%
  print()

```

##### Summary statistics of title_ratings

```{r specific-summary-stat-2, echo=FALSE, warning=FALSE, message=FALSE}

title_ratings_summary <- title_ratings %>%
  summarise(num_rows=n(),
            num_cols=ncol(.),
            missing_values_tconst = sum(is.na(tconst)),
            missing_values_averageRating = sum(is.na(averageRating)),
            missing_values_numVotes = sum(is.na(numVotes))
  ) %>%
  print()

```

##### Summary statistics of title_basics

```{r specific-summary-stat-3, echo=FALSE, warning=FALSE, message=FALSE}

title_basics_summary <- title_basics %>%
  summarise(num_rows=n(),
            num_cols=ncol(.),
            missing_values_tconst = sum(is.na(tconst)),
            missing_values_titleType = sum(is.na(titleType)),
            missing_values_startYear= sum(is.na(startYear)),
            missing_values_endYear= sum(is.na(endYear))
  )%>%
  print()

```

```{r merging, echo=FALSE, warning=FALSE, message=FALSE}
# filtering TV series alone using titleType(tvEpisode) from title_basics

tv_series <- title_basics %>%
  filter(titleType == "tvEpisode")

# Merged with ratings using tconst

tv_series_ratings <- tv_series %>%
  inner_join(title_ratings, by = "tconst")
print(nrow(tv_series_ratings))

# Merging with title_episode using tconst to add parentTconst to tv_series_ratings

tv_series_ratings_with_parent <- tv_series_ratings %>%
  left_join(title_episode %>% 
              select(tconst,parentTconst), by= "tconst")
head(tv_series_ratings_with_parent)
View(tv_series_ratings_with_parent)
```

### Visualizations

#### TV Series Length vs Average Rating

We will also create some visualizations to better understand the datasets.

We plot for Series Length vs Average Rating. To Compute Series Length we calculate the length of each TV series based on the number of episodes and the number of years the series was telecast-ed.

####  Plotting TV Series Length (total episodes) vs Average Ratings.

```{r series-length-vs-rating-episodes, echo=FALSE, warning=FALSE, message=FALSE}

# Compute the number of episodes for each series

total_episodes <- title_episode %>%
  group_by(parentTconst) %>%
  summarize(total_episodes = n(), .groups = 'drop')  # Count episodes

# Compute and sort the number of episodes for each series
total_episodes_sorted <- title_episode %>%
  group_by(parentTconst) %>%
  summarize(total_episodes = n(), .groups = 'drop') %>%
  arrange(desc(total_episodes))

# Select and rename the columns from tv_series_ratings_with_parent
ratings_to_merge <- tv_series_ratings_with_parent[, c("parentTconst", "averageRating")]
names(ratings_to_merge)[names(ratings_to_merge) == "averageRating"] <- "averageRating_series_wise"

# Merge the two data frames
total_episodes_sorted <- merge(total_episodes_sorted, 
                               ratings_to_merge, 
                               by = "parentTconst", 
                               all.x = TRUE)

# scatter plot 
library(ggplot2)


ggplot(total_episodes_sorted, aes(x = total_episodes, y = averageRating_series_wise)) +
  geom_jitter(na.rm = TRUE, alpha = 0.3, size = 1, width = 0.1) +  # Add jitter
  labs(
    title = "Relationship between Total Episodes and Average Customer Ratings",
    x = "Total number of Episodes",
    y = "Average customer Rating"
  ) +
  theme_minimal()
```


#### Plotting TV Series Length (total years run) vs Average Ratings.

```{r total-years-vs-rating, echo=FALSE, warning=FALSE, message=FALSE}

# Convert endYear to character if it is numeric
tv_series_ratings_with_parent <- tv_series_ratings_with_parent %>%
  mutate(
    endYear = as.character(endYear),  
    endYear = na_if(endYear, '\\N'),  
    endYear = as.numeric(endYear),    
    startYear = as.numeric(startYear),
    length = ifelse(is.na(endYear), 2024 - startYear, endYear - startYear) # Calculate length
  )

# Scatter plot for Length of TV Series (years) vs Average Ratings

ggplot(tv_series_ratings_with_parent, aes(x = length, y = averageRating)) +
  geom_jitter(alpha = 0.2, size = 0.7, width = 0.1, height = 0.1) +  # Further reduce point size and opacity
  geom_smooth(method = "lm", color = "red", se = FALSE) +  # Add a linear trend line
  labs(
    x = "Length of TV Series (years)",
    y = "Average Rating",
    title = "Length of TV Series vs. Average Rating"
  ) +
  theme_minimal(base_size = 15) +  # Increase text size for better readability
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Center the title and make it bold
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  )
```


## Data Preparation

#### Counting the number of missing values (NA) per dataset 

```{r counting_NA, echo=FALSE, warning=FALSE, message=FALSE}
#Amount of NA's of title_basics
Sum_NA_title_basics <-sum(is.na(title_basics))

#Amount of NA's of title_episode
Sum_NA_title_episode <- sum(is.na(title_episode))

#Amount of NA's of title_ratings
Sum_NA_title_ratings <- sum(is.na(title_ratings))

#We can see that the title_ratings dataset has no NA's. 
```


### Motivation for dropping missing values (NA)

#### title_basics:

After examining the title_basics dataset, we noticed a significant number of missing values (NAs) in the columns "StartYear," "EndYear," "runtimeMinutes," and "genre." Since our project requires calculating the series length (series_length = EndYear - StartYear), it is crucial to remove rows with missing values in the "StartYear" and "EndYear" columns. Retaining rows with NAs in these columns would result in inaccurate calculations.

Additionally, the columns "runtimeMinutes" and "genre" are not necessary for our analysis, so we will drop them.

Regarding the missing values in "EndYear," there is no feasible way to recover this data in our current dataset. Because the "StartYear" and "EndYear" are specific to each TV series, it does not make sense to impute the missing values with estimates or averages, as this would compromise the accuracy of our analysis. Therefore, we will drop all rows where either "StartYear" or "EndYear" is missing to ensure the integrity of our data.


#### title_episode:

In this dataset, we observe missing values (NAs) in the columns "seasonNumber" and "episodeNumber." Since this information is specific to each TV series, we cannot impute these missing values using data from other TV shows, as doing so would introduce inaccuracies. Additionally, the missing values in these columns are missing at random. Therefore, it is important to drop rows with missing values in these columns.


#### title_ratings:

This dataset does not contain missing values(NA) 


### Revised summary statistics after dropping the missing values : title_basics 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
title_basics_no_NAs <- drop_na(title_basics)
View(title_basics_no_NAs)
summary(title_basics_no_NAs)
```


### Revised summary statistics after dropping the missing values : title_episode 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
title_episode_no_NAs <- drop_na(title_episode)
View(title_episode_no_NAs)
summary(title_episode_no_NAs) 
```


### Merging the datsets

```{r, echo=FALSE, warning=FALSE, message=FALSE}
merged_data <- title_basics_no_NAs %>%
  inner_join(title_ratings, by = "tconst") %>%
  inner_join(title_episode_no_NAs, by = c("tconst" = "parentTconst"))
```


### Engineering for new variables: 

#### total_years

A new variable is created that shows how many years a series has run, called total_years. This variable will be one of the independent variable in our analysis, and is created by deducting the first year that the series ran from the last year. 


#### episode_count

A new variable is created that shows the total number of episodes a series has, called episode_count. This variable will be one of the independent variable in our analysis.

```{r variable engineering, echo=FALSE,warning=FALSE, message=FALSE}
merged_data <- merged_data %>%
  mutate(total_years = endYear - startYear)


# Calculate the total number of episodes for each series
episode_count <- title_episode_no_NAs %>%
  group_by(parentTconst) %>%
  summarise(episode_count = n(), .groups = 'drop')  # Count the number of episodes per series


# Merge the episode count back into the merged_data
merged_data <- merged_data %>%
  left_join(episode_count, by = c("tconst" = "parentTconst"))

```


### Removing outliers 

Using the Interquartile Range (IQR) method, we detect outliers in the variables considered for analysis by calculating the spread of the data.The IQR measures the spread of the middle 50% data, making the data robust for analysis by handling the skewed data.This is the last step in preparing the data for analysis.


```{r}
remove_outliers_iqr <- function(column) {
  Q1 <- quantile(column, 0.25)  # Calculate the 25th percentile
  Q3 <- quantile(column, 0.75)  # Calculate the 75th percentile
  IQR_value <- IQR(column)  # Calculate the IQR (Q3 - Q1)
  lower_bound <- Q1 - 1.5 * IQR_value  # Lower bound for outliers
  upper_bound <- Q3 + 1.5 * IQR_value  # Upper bound for outliers
  column >= lower_bound & column <= upper_bound # Returns TRUE for values within bounds
}

# Columns to check for outliers
columns_to_check <- c("total_years", "episode_count", "numVotes", "averageRating")

# Apply outlier removal and combine results
cleaned_data <- merged_data[
  rowSums(sapply(merged_data[columns_to_check], remove_outliers_iqr)) == length(columns_to_check), 
]

```


