---
title: "Data Exploration"
author: "team 03"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message= FALSE)
```


### Install necessary pacakges for data exploration

The following libraries are required for this analysis: 

- **readr**: For reading data files.

- **tidyverse**: For data manipulation and analysis.

- **dplyr**: For data wrangling and manipulation.

- **ggplot2**: For data visualization.


```{r install packages,include=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Introduction

This report provides the initial data exploration of the data sets. Our goal is to provide the structure of our data set,summarize key variables,and provide some basic exploratory visualizations. The data set contains information about TV series, including customer ratings, number of episodes and series duration.
 
## Overview of the Dataset and Variables

The dataset used in the project contains information about TV series and episodes from IMD Since we are interested in understanding the relationship between the length of the TV series and it's average customer ratings, we focus on the three datasets listed and described below.

- **title.episode.tsv.gz**: Contains information about individual episodes of TV series.

- **title.basics.tsv.gz**: Contains information about TV series titles,the start year,genres,etc.

- **title.ratings.tsv.gz**: Contains user ratings for titles.




## Description of Variables

#### title.episode.tsv.gz

```{r variables-episode,echo=FALSE}
variables_episode <- data.frame(
  Variables = c("`tconst`","`parentTconst`","`seasonNumber`", "`episodeNumber`"),
  Description = c(
    "An alphanumeric identifier unique to each episode.",
    "An alphanumeric identifier for the parent TV series of the episode.It links the episode to the overall series.",
    "The season number that the episode belongs to within the TV series.",
    "The specific episode number of the `tconst` in the TV series."
                  ),
  Class = c("String","String","Interger","Interger")
)

# Display the table using knitr::kable()
knitr::kable(variables_episode,col.names = c("Variables","Description","Class"))
```

#### title.basics.tsv.gz

```{r variables-basics,echo=FALSE}
# Create a data frame for title.basics.tsv.gz
variables_basics <- data.frame(
  Variables = c("`startYear`", "`endYear`"),
  Description = c(
    "The year the series began.",
    "TV Series end year."
                 ),
  Class = c("Integer", "Integer")
)

# Display the table using knitr::kable()
knitr::kable(variables_basics, col.names = c("Variables", "Description","Class"))

```

#### title.ratings.tsv.gz

```{r variables-ratings, echo=FALSE}
# Create a data frame for title.ratings.tsv.gz
variables_ratings <- data.frame(
  Variables = c("`tconst`", "`averageRating`"),
  Description = c(
    "Alphanumeric unique identifier of the title.",
    "Weighted average of all individual user ratings."
  ),
  Class = c("String", "Numeric")
)

# Display the table using knitr::kable()
knitr::kable(variables_ratings, col.names = c("Variables", "Description", "Class"))

```


### Loading the Datasets

Programmatically loading the following datasets from IMDb:

- **title.basics.tsv.gz**: Contains information about movie and TV series titles, their start year, genres, etc.

- **title.episode.tsv.gz**: Contains information about individual episodes of TV series.

- **title.ratings.tsv.gz**: Contains user ratings for titles.


```{r load-datasets, echo=FALSE, warning=FALSE, message=FALSE}

# URLs to the IMDb datasets
urls <- c(
  "https://datasets.imdbws.com/title.basics.tsv.gz",
  "https://datasets.imdbws.com/title.episode.tsv.gz",
  "https://datasets.imdbws.com/title.ratings.tsv.gz"
)

# Load the datasets
datasets <- lapply(urls, read_delim, delim = '\t', na = '\\N')

# Access individual datasets
title_basics <- datasets[[1]]
title_episode <- datasets[[2]]
title_ratings <- datasets[[3]]

```

### Initial Data Exploration

#### Summary Statistics 

##### Get a quick overview of the data by using the code summary()

###### Summary statistics: title_basics


```{r summary-stats, echo=FALSE, warning=FALSE, message=FALSE}
# Summary statistics for each dataset
summary(title_basics)
```


###### Summary statistics: title_episode

```{r}
summary(title_episode)
```


###### Summary statistics: title_ratings

```{r}
summary(title_ratings)
```


#### Specific summary statistics

###### Summary statistics : title_episode

```{r specific-summary-stats, echo=FALSE, warning=FALSE, message=FALSE}

title_episode_summary <- title_episode %>%
  summarise(num_rows=n(),
            num_cols=ncol(.),
            missing_values_tconst=sum(is.na(tconst)),
            missing_values_parentTconst=sum(is.na(parentTconst)),
            missing_values_seasonNumber=sum(is.na(seasonNumber)),
            missing_values_episodeNumber=sum(is.na(episodeNumber)) 
  )%>%
  print()

```

##### Summary statistics : title_ratings

```{r specific-summary-stat-2, echo=FALSE, warning=FALSE, message=FALSE}

title_ratings_summary <- title_ratings %>%
  summarise(num_rows=n(),
            num_cols=ncol(.),
            missing_values_tconst = sum(is.na(tconst)),
            missing_values_averageRating = sum(is.na(averageRating)),
            missing_values_numVotes = sum(is.na(numVotes))
  ) %>%
  print()

```

##### Summary statistics : title_basics

```{r specific-summary-stat-3, echo=FALSE, warning=FALSE, message=FALSE}

title_basics_summary <- title_basics %>%
  summarise(num_rows=n(),
            num_cols=ncol(.),
            missing_values_tconst = sum(is.na(tconst)),
            missing_values_titleType = sum(is.na(titleType)),
            missing_values_startYear= sum(is.na(startYear)),
            missing_values_endYear= sum(is.na(endYear))
  )%>%
  print()

```

```{r merging, echo=FALSE, warning=FALSE, message=FALSE}
# filtering TV series alone using titleType(tvEpisode) from title_basics

tv_series <- title_basics %>%
  filter(titleType == "tvEpisode")

# Merged with ratings using tconst

tv_series_ratings <- tv_series %>%
  inner_join(title_ratings, by = "tconst")


# Merging with title_episode using tconst to add parentTconst to tv_series_ratings

tv_series_ratings_with_parent <- tv_series_ratings %>%
  left_join(title_episode %>% 
              select(tconst,parentTconst), by= "tconst")

```

### Visualizations

#### TV Series Length vs Average Rating

To understand the relationship between the length of TV series and average customer ratings, we will create visualizations..This help us explore trends and patterns in the data.

The length of the TV series will be determined based on the number of episodes or the number of years the series was aired.


#### Plot 1: Total Episodes vs. Average Rating

```{r series-length-vs-rating-episodes, echo=FALSE, warning=FALSE, message=FALSE}

# Compute the number of episodes for each series

total_episodes <- title_episode %>%
  group_by(parentTconst) %>%
  summarize(total_episodes = n(), .groups = 'drop')  # Count episodes

# Compute and sort the number of episodes for each series
total_episodes_sorted <- title_episode %>%
  group_by(parentTconst) %>%
  summarize(total_episodes = n(), .groups = 'drop') %>%
  arrange(desc(total_episodes))

# Select and rename the columns from tv_series_ratings_with_parent
ratings_to_merge <- tv_series_ratings_with_parent[, c("parentTconst", "averageRating")]
names(ratings_to_merge)[names(ratings_to_merge) == "averageRating"] <- "averageRating_series_wise"

# Merge the two data frames
total_episodes_sorted <- merge(total_episodes_sorted, 
                               ratings_to_merge, 
                               by = "parentTconst", 
                               all.x = TRUE)

# scatter plot 
library(ggplot2)


ggplot(total_episodes_sorted, aes(x = total_episodes, y = averageRating_series_wise)) +
  geom_jitter(na.rm = TRUE, alpha = 0.3, size = 1, width = 0.1) +  # Add jitter
  labs(
    title = "Relationship between Total Episodes and Average Customer Ratings",
    x = "Total number of Episodes",
    y = "Average customer Rating"
  ) +
  theme_minimal()
```


##### **Inference :**



#### Plot 2: Total number of years the series was aired vs. Average Rating


```{r total-years-vs-rating, echo=FALSE, warning=FALSE, message=FALSE}

# Convert endYear to character if it is numeric
tv_series_ratings_with_parent <- tv_series_ratings_with_parent %>%
  mutate(
    endYear = as.character(endYear),  
    endYear = na_if(endYear, '\\N'),  
    endYear = as.numeric(endYear),    
    startYear = as.numeric(startYear),
    length = ifelse(is.na(endYear), 2024 - startYear, endYear - startYear) # Calculate length
  )

# Scatter plot for Length of TV Series (years) vs Average Ratings

ggplot(tv_series_ratings_with_parent, aes(x = length, y = averageRating)) +
  geom_jitter(alpha = 0.2, size = 0.7, width = 0.1, height = 0.1) +  # Further reduce point size and opacity
  geom_smooth(method = "lm", color = "red", se = FALSE) +  # Add a linear trend line
  labs(
    x = "Length of TV Series (years)",
    y = "Average Rating",
    title = "Length of TV Series vs. Average Rating"
  ) +
  theme_minimal(base_size = 15) +  # Increase text size for better readability
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Center the title and make it bold
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  )
```


##### **Inference :**


## Data Preparation

### Remove NAs from title_basics dataset and show the new summary statistics

```{r, echo=FALSE}
title_basics_no_NAs <- drop_na(title_basics)
View(title_basics_no_NAs)
summary(title_basics_no_NAs)
```

### Remove NAs from title_episode dataset and show the new summary statistics
```{r, echo=FALSE}
title_episode_no_NAs <- drop_na(title_episode)
View(title_episode_no_NAs)
summary(title_episode_no_NAs) 
```

### Remove NAs from title_ratings dataset and show the new summary statistics
```{r, echo=FALSE}
title_ratings_no_NAs <- drop_na(title_ratings)
View(title_ratings_no_NAs)
summary(title_ratings_no_NAs) 
```

### Merge the three datsets together
```{r, echo=FALSE}
merged_data <- title_basics_no_NAs %>%
  inner_join(title_ratings_no_NAs, by = "tconst") %>%
  inner_join(title_episode_no_NAs, by = c("tconst" = "parentTconst"))
```

### Create new variabel: series_length
A new variable is created that shows how many years a series has run, called series_length. This variable will be the independent variable in our analysis, and is created by deducting the first year that the series ran from the last year. 
```{r, echo=FALSE}
merged_data <- merged_data %>%
  mutate(series_lengths = endYear - startYear)
```



