---
title: "Regression: Assumptions & Conclusion"

author: "team 03"
date: "2024-09-27"
output: html_document
---

## 1. Assumptions of Linear Regression

The assumptions of the linear regression models are checked by generating diagnostic plots.The residual plots were first run with only log transformations for `numVotes` in model1 and log transformations for `episode_count` and `numVotes` in model2. However, later the `averageRating` was also log transformed to stabilize variance.

```{r}
knitr::opts_chunk$set( include=FALSE,echo = FALSE, warning = FALSE, message= FALSE,fig.width = 8,fig.height = 6,fig.align = 'center',dpi=300,dev='png',fig.retina = 2)
```


```{r setup, include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
library(ggplot2)
library(car)
library(dplyr)
```


```{r,include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
# Load the regression models
model1 <- readRDS("../../gen/output/model1.rds")
model2 <- readRDS("../../gen/output/model2.rds")
```


### 1.2. Model 1 : total_years


### (a). Residuals vs. Fitted Plot

```{r residuals plot model 1,include=TRUE,warning=FALSE,echo=FALSE,message=FALSE,fig.cap=NULL}
plot(model1, which = 1)
```

#### Inference:

Residuals vs Fitted plot helps to visually assess whether the assumptions of homoscedasticity hold for model1. The residuals represent the difference between the observed values of `averageRating` (log-transformed) and the fitted values from the model.The residuals appear to be mostly scattered without a visible trend, suggesting that the assumption of linearity has been met.There is clustering visible on the right side, but the spread is less pronounced compared to before log transformation, indicating a positive impact on stabilizing variance.The reduction indicates that the issue of homoscedasticity has largely improved.But this has been acknowledged as  a limitation and we further proceed with the model1.

### (b). Q-Q Plot for normality check

```{r qqplot model 1,include=TRUE,warning=FALSE,echo=FALSE,message=FALSE,fig.cap=NULL}
plot(model1, which = 2)

```

#### Inference:

The Q-Q plot assesses whether the residuals from the regression model follow a normal distribution. Most residuals are generally aligned along the diagonal, indicating that the model captures a reasonable level of normality for the residuals. However, there are noticeable deviations at both extremes of the distribution. This suggests that there may be some non-normality in the extreme values.Given the slight deviation we continue to proceed with the current model since we have already log transformed the variables.


### (c). Scale-Location Plot for homoscedasticity check

```{r scale-loc plot model 1,include=TRUE,warning=FALSE,echo=FALSE,message=FALSE,fig.cap=NULL}
plot(model1, which = 3)
```

#### Inference:

The scale-location plot provides a visual assessment of the homoscedasticity of residuals.There appears to be increasing variance in the residuals as the fitted values increase. This is a potential limitation, as non-constant variance may lead to unreliable standard errors, which could affect the interpretation of the results.Nevertheless, for the purposes of this analysis, we have decided to proceed with the current model without making further adjustments.But we acknowledge this as a limitation. Future analyses may consider transformations or alternative regression techniques to address this issue and improve model assumptions. 


### (d). Residuals vs. Leverage Plot to detect influential points

```{r residual leverage model 1,include=TRUE,warning=FALSE,echo=FALSE,message=FALSE,fig.cap=NULL}
plot(model1, which = 5)
```


#### Inference:

Influential points with high leverage remain even after the IQR-based outlier removal, this suggests that the dataset still has few influential observations that could be affecting the model's stability and predictions. Given these are few in number we decide to overlook it and proceed with our analysis.Note,that these influential points may disproportionately impact the model's predictions. 


### 1.3. Model 2: log_episode_count

### (a). Residuals vs. Fitted Plot


```{r residuals plot model 2,include=TRUE,warning=FALSE,echo=FALSE,message=FALSE,fig.cap=NULL}

plot(model2, which = 1)

```

#### Inference:

The residuals are mostly scattered around the zero line, without a visible trend, suggesting that the assumption of linearity has been met.Some clustering visible may indicate minor deviations.The residual spread appears relatively uniform, although there are denser clusters of points, especially towards the right side of the plot. This could indicate some slight non-constant variance in the residuals.There are no major signs of strong funnel shapes or systematic patterns that would suggest significant heteroscedasticity.


### (b). Q-Q Plot for normality check

```{r qqplot model 2,include=TRUE,warning=FALSE,echo=FALSE,message=FALSE,fig.cap=NULL}
plot(model2, which = 2)

```

#### Inference:

The deviations at the tails imply that the model does not fully satisfy the normality assumption, especially in extreme values. The residuals appear roughly normal in the middle section.We acknowledge this as a limitation but decide to proceed with the current model, documenting that the tails' departure from normality might impact statistical inference.


### (c). Scale-Location Plot for homoscedasticity check

```{r scale-loc plot 2,include=TRUE,warning=FALSE,echo=FALSE,message=FALSE,fig.cap=NULL}
plot(model2, which = 3)
```


#### Inference:

There is a discernible spread in the residuals and a pattern exists implying the homoscedasticity assumption does not hold perfectly.Since the deviatins are not extreme we acknowledge this limitation and highlight that the precision of the model maybe reduced.


### (d). Residuals vs. Leverage Plot to detect influential points

```{r leverage 2,include=TRUE,warning=FALSE,echo=FALSE,message=FALSE,fig.cap=NULL}
plot(model2, which = 5)

```


#### Inference:

The plot suggests that most of the data points are not highly influential, but there are some points with slightly higher leverage. Since tehre are no extreme influencers,we proceed with the model while noting these few potentially influential points as a minor limitation.



## 2. Regression output inference

## 2.1. Model 1 : `total_years`

### (a). Summary:

```{r summary-model1,warning=FALSE,echo=FALSE,message=FALSE}
summary(model1)
```

### (b). Inference: 

### Statistical Significance: 

All the p-values are less than 0.05,indicating evidence against the null hypothesis suggesting that both `total_years` and `log_numVotes` have statistically significant effect on `log_averageRating`

### Intercept: 

The intercept estimate of 2.107e+00 represents the expected `log_averageRating` when both `total_years` and `log_numVotes` are zero.This serves as a baseline for calculating expected ratings

### `total_years` (-0.0023): 

This translates to approximately a 0.23% decrease in average rating per additional year holding `numVotes` constant. This indicates that longer-running series tend to experience a slight decline in average ratings, possibly due to factors like viewer fatigue or decreasing quality of teh show over time.

### `log_numVotes` (-0.0109): 

The coefficient for `log_numVotes` is `-0.0109`. Since `numVotes` is log-transformed,for a 1% increase in the number of votes, the average rating decreases by  0.0109%. The negative relationship suggests that as more audience rates a show, the average rating tends to slightly decrease, possibly because of critical reviews.

### Model Fit : (R2=0.01443)

A R2 value of 0.0144 indicates that the model has a low explanatory power and only a small portion of the variance in `log_averageRating` is explained by the model.It implies there maybe other factors explaining the variance in the dependent variable.


## 2.2. Model 2 : `episode_count`

### (a). Summary:

```{r summary-model2,warning=FALSE,echo=FALSE,message=FALSE}
summary(model2)
```

### (b).  Inference: 

### Statistical Significance: 

Both `log_episode_count` and `log_numVotes` have p values lower than 0.05 (p-values < 2e-16), suggesting that their effects on `log_averageRating` are statistically significant.

### `log_episode_count` (-0.01379)

A 1% increase in the number of episodes (`episode_count`) is associated with a 0.01379% decrease in the average rating (`averageRating`), holding the number of votes constant.

### ` log_numVotes` (-0.01169)

A 1% increase in the number of votes (`numVotes`) is associated with a 0.01169% decrease in the average rating. This suggests that more popular shows (in terms of the number of votes) tend to receive slightly lower ratings, possibly because a larger audience and varied opinion.

### Model Fit : (R2= 0.02425)

2.43% of the variation in `log_averageRating` is explained by the number of episodes and the number of votes.It implies there maybe other factors explaining the variance in the dependent variable.
 

## 3. Conclusion

The analysis findings suggest that both `total_years` and `episode_count` have statistical significance but small negative effects on the average ratings. As the length of a TV series, both in terms of number of episodes or the number of years it has aired—increases, there is a slight decline in viewer ratings. The control variable, `numVotes`,showed a small negative impact, indicating that higher popularity might lead to more critical reviews.The analysis also suggests that while the length of a series does have an impact on ratings, it is likely not the only or even the primary factor affecting viewer satisfaction. While longer series may offer many benefits, including deep character development, it may also risk losing viewers' interest over time.Further research could include exploring additional variables,to better understand the multifaceted nature of viewer satisfaction.


